---
title: "Preprocessing"
author: "Polina Yarko"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_document
---
#Initial Preprocessing
This script performs the initial preprocessing by merging all participant-level CSV files into a single dataset, cleans the transcribed responses, calculating several string distance measures, and classifying the responses based on those string distances. SPQ scores are also added to the dataset.
```{r setup}
library(jsonlite)
library(dplyr)
library(stringr)
library(readr)
library(purrr)
library (stringdist)
library(here) #this R-Script should be in your preprocessing folder
library(readxl)
```

## Load and merge participant-level CSV files 

```{r restructure study result file}
# ---------------------------- 
# 1. find & merge study result files from all participants
# ---------------------------- 
# get the study_result_... csv-files
study_files <- list.files( 
  here("demo_results", "participant_exports"), 
  pattern = "^study_result_.*\\.csv$", 
  full.names = TRUE 
) 
#merge study results from all participants
all_data <- purrr::map_dfr(study_files, function(file) { 
read.csv(file) %>% 
  mutate( 
    recording_key = stringr::str_remove(file_name, "^response_") %>%
                    stringr::str_remove("\\.txt$"), 
    study_file = basename(file) 
  ) 
}) 

# ---------------------------- 
# 2. restructure document 
# ---------------------------- 
final_data <- all_data %>%
  relocate(Sentence, Answer, .before = mic_input_text) %>%
  relocate(cloze, .after = mic_input_text) %>%
  relocate(study_file, .after = participant_id) %>%
  select(participant_id, study_file, sentence_index, everything())
final_data <- final_data %>%
  select(-round_index, -file_name)
```

## Preprocess transcribed responses
### Clean and standardize transcribed responses

The transcribed responses are converted to lowercase and and punctuation is removed to standardize responses for string distance calculations.
The original transcriptions are preserved in a separate column before preprocessing. 

```{r convert to lowercase & remove punctuation}
# duplicate transcribed responses column
final_data$mic_input_text_raw <- final_data$mic_input_text
final_data <- final_data %>% relocate(mic_input_text_raw, .after = Sentence)

# convert to lowercase
final_data$mic_input_text<- tolower(final_data$mic_input_text)
# remove puctuation
final_data$mic_input_text<- str_remove_all(final_data$mic_input_text, "[[:punct:]]") 

# create vectors for string distance
string_vec1<- final_data$Answer
string_vec2<- final_data$mic_input_text
```

Responses containing "nothing" are retained as valid responses. 
Responses containing numbers are excluded. 
For the remaining responses, common function words are removed and repeated-word responses are reduced to the repeated word. 

```{r convert numbers to NA & extract last word in case of repetition}
final_data <- final_data %>%
  mutate(
    mic_input_text_clean = case_when(
      str_detect(mic_input_text, "nothing") ~ "nothing",
      str_detect(mic_input_text, "\\d") ~ NA_character_,
      TRUE ~ { mic_input_text %>%
        str_remove_all("\\bthe\\b|\\ba\\b|\\band\\b") %>%
        str_squish() %>%
        sapply(function(x) {
          words <- str_split(x, " ")[[1]]
          dupes <- words[duplicated(words)]
          if (length(dupes) == 0) x else unique(dupes) %>% str_c(collapse = " ")
        }, USE.NAMES = FALSE) 
      }
    )
  )
final_data <- final_data %>%
  dplyr::relocate(mic_input_text_clean, .after = mic_input_text)
```

Responses that still contain more than one word are excluded from further analysis.

```{r clean multiple-word responses}
# check if there is more than 1 word
final_data$multiple_words <- ifelse(
  stringr::str_count(
    stringr::str_trim(final_data$mic_input_text_clean),
    "\\S+"
  ) > 1,
  1,
  0
)
final_data <- final_data %>%
  dplyr::relocate(multiple_words, .after = mic_input_text_clean)
# replace multiple strings with NA
final_data <- final_data %>%
  mutate(
    mic_input_text_clean = if_else(
      multiple_words == 1,
      NA_character_,
      mic_input_text_clean
    )
  )
```

### Compute string distances

Different string distance measures are calculated to determine which measure best assesses the similarity between the target word and the transcribed response for each sentence.

```{r compute string distances}
final_data$levenshtein <- stringdist(string_vec1, string_vec2, method = 'lv')
final_data$optimal_string_alignment<- stringdist(string_vec1, string_vec2, method = 'osa')
final_data$damerau_levenshtein<- stringdist(string_vec1, string_vec2, method = 'dl') 
final_data$longest_common_string<- stringdist(string_vec1, string_vec2, method = 'lcs')
final_data$jaro_winkler<- stringdist(string_vec1, string_vec2, method='jw', p=0.1)
final_data$soundex<- stringdist(string_vec1, string_vec2, method = 'soundex')

final_data <- final_data %>%
  relocate(
    levenshtein,
    optimal_string_alignment,
    damerau_levenshtein,
    longest_common_string,
    jaro_winkler,
    soundex,
    .after = multiple_words
  )
```

Levenshtein distance and Jaro-Winkler distance showed the best fit for classifying the transcribed responses. 

The classification for the Levenshtein distance was defined as follows:
- noword: responses containing "nothing"
- correctword: Levenshtein distance ≤ 1
- word: Levenshtein distance ≥ 2

```{r word_noword_levenshtein column}
final_data <- final_data %>%
  mutate(
    word_noword_levenshtein = case_when(
      str_detect(mic_input_text_clean, "nothing") ~ "noword",
      levenshtein <= 1 ~ "correctword",
      levenshtein >= 2 ~ "word",
      TRUE ~ NA_character_
    )
  )

final_data <- final_data %>%
  relocate(word_noword_levenshtein, .after = multiple_words)
```

The classification for the Jaro-Winkler distance was defined as follows:
- noword: responses containing "nothing"
- correctword: Jaro-Winkler distance ≤ 0.2
- word: Jaro-Winkler distance > 0.2

The Jaro-Winkler distance values were used for subsequent analyses.
```{r word_noword_jw column}
final_data <- final_data %>%
  mutate(
    word_noword_jw = case_when(
      str_detect(mic_input_text_clean, "nothing") ~ "noword",
      jaro_winkler <= 0.2 ~ "correctword",
     TRUE ~ "word"
    )
  )

final_data <- final_data %>%
  relocate(word_noword_jw, .after = (word_noword_levenshtein))
```

## Add SPQ scores

SPQ scores are added to the dataset for subsequent analysis.

```{r add SPQ scores}
# load SPQ data
df_spq <- read_excel("spq_data.xlsx")

# merge with final_data
final_data <- left_join(final_data, df_spq, by = "participant_id")
```

## Save dataset as CSV

```{r save as csv}
write.csv(final_data, paste0("final_data_", format(Sys.Date(), "%m%d%y"), ".csv"), row.names = FALSE)
```