---
title: "Response Proportions"
output: html_document
---
# Prepare data for modelling
This script takes the preprocessed dataset generated by '01_preprocessing_PLang_EN.Rmd' and prepares it for modelling. 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(tidyverse)
library(dplyr)
library(grid)
library(gridExtra)
```

## Load data

```{r load-data}
dat <- read_csv(".../preprocessing/final_data_MMDDYY.csv") #your path to final dataset generated by '01_preprocessing_PLang_EN.Rmd'
#%>%
#  mutate(
#    condition = case_when(
#      str_detect(file_name, "el_mm") ~ "low_entropy_mm",
#      str_detect(file_name, "el_") ~ "low_entropy",
#      str_detect(file_name, "em_hc") ~ "medium_entropy",
#      str_detect(file_name, "em_lc") ~ "medium_entropy_mm",
#      str_detect(file_name, "eh") ~ "high_entropy",
#      TRUE ~ NA_character_
#    ),
#    condition = factor(condition,levels = c("low_entropy","medium_entropy","high_entropy","medium_entropy_mm","low_entropy_mm"))
#  )
dat$condition<-as.factor(dat$condition)
dat$condition<-factor(dat$condition,levels = c("el","em_hc","eh","el_mm","em_lc"))
dat <- dat[!is.na(dat$mic_input_text_clean),]
 
dat<-dat%>%
  mutate(
      block = case_when(
      str_detect(randomisation, "Mini") ~ "mini",
      str_detect(randomisation, "Midi") ~ "midi",
      str_detect(randomisation, "Maxi") ~ "maxi",
      TRUE ~ NA_character_
    ),
    block = factor(block, levels = c("mini","midi","maxi"))
  )


dat<-dat%>%
  mutate(
      channels = case_when(
      str_detect(recording_key, "tw_12") ~ "12",
      str_detect(recording_key, "tw_3") ~ "3",
      str_detect(recording_key, "tw_6") ~ "6",
      str_detect(recording_key, "tw_1") ~ "1",
      TRUE ~ NA_character_),
    channels = factor(channels, levels = c("1","3","6","12"))
  )


dat
response_levels <- c("noword", "word", "correctword")

#dat$word_noword_jw<-NA

#dat <- dat %>%
#  mutate(
#    word_noword_jw = case_when(
#      str_detect(mic_input_text, "nothing") ~ "noword",
#      str_detect(mic_input_text, "\\s+") ~ "noword",  # <- more than one word
#      jaro_winkler <= 0.2 ~ "correctword",
#      is.na(word_noword_jw) ~ "word",
#      TRUE ~ NA_character_
#    )
#  )

dat$clarity_rating<-dat$clarity_rating/100
dat$confidence_rating<-dat$confidence_rating/100

colnames(dat)

randomisation_list <- dat %>%
  group_by(participant_id) %>%
  summarise(
    unique_randomisation = paste(unique(randomisation), collapse = ", "),
    .groups = "drop"
  )
```

## Calculate subject-level proportions
Calculate the proportion of each response category ('word', 'correctword', and 'noword') for each participant, separately for the Levenshtein and Jaro-Winkler classifications.

```{r calculate-proportions}
prop_levenshtein <- dat %>%
  filter(!is.na(word_noword_levenshtein)) %>%
  count(participant_id, word_noword_levenshtein, name = "n") %>%
  group_by(participant_id) %>%
  mutate(proportion = n / sum(n)) %>%
  ungroup() %>%
  complete(
    participant_id,
    word_noword_levenshtein = response_levels,
    fill = list(n = 0, proportion = 0)
  ) %>%
  mutate(word_noword_levenshtein = factor(word_noword_levenshtein, levels = response_levels))

prop_jw <- dat %>%
  filter(!is.na(word_noword_jw)) %>%
  count(participant_id, word_noword_jw, name = "n") %>%
  group_by(participant_id) %>%
  mutate(proportion = n / sum(n)) %>%
  ungroup() %>%
  complete(
    participant_id,
    word_noword_jw = response_levels,
    fill = list(n = 0, proportion = 0)
  ) %>%
  mutate(word_noword_jw = factor(word_noword_jw, levels = response_levels))
```

## Plot: word_noword_levenshtein

```{r}
ggplot(prop_levenshtein, aes(x = word_noword_levenshtein, y = proportion)) +
  geom_boxplot(outlier.shape = NA, width = 0.55) +
  geom_point(aes(group = participant_id), position = position_jitter(width = 0.12, height = 0), alpha = 0.65, size = 2) +
  scale_y_continuous(limits = c(0, 1), labels = scales::percent_format(accuracy = 1)) +
  labs(
    x = NULL,
    y = "Proportion of responses Levenshtein"
  ) +
  theme_classic(base_size = 14)
```

## Plot: word_noword_jw

```{r}
ggplot(prop_jw, aes(x = word_noword_jw, y = proportion)) +
  geom_boxplot(outlier.shape = NA, width = 0.55) +
  geom_point(aes(group = participant_id), position = position_jitter(width = 0.12, height = 0), alpha = 0.65, size = 2) +
  scale_y_continuous(limits = c(0, 1), labels = scales::percent_format(accuracy = 1)) +
  labs(
    x = NULL,
    y = "Proportion of responses JW"
  ) +
  theme_classic(base_size = 14)
```

## Proportions by conditions
Calculate the proportion of each response category ('word', 'correctword', and 'noword') by experimental conditions ('el_mm', 'el', 'em_hc', 'em_lc', and 'eh') and trial blocks ('midi', 'mini', and 'maxi'), separately for the Levenshtein and Jaro-Winkler classifications.
Note: Proportions are calculated separately for each block (i.e., mini = 60 trials, midi = 60 trials, maxi = 80 trials).

```{r calculate-proportions-cond}
prop_levenshtein_cond <- dat %>%
  filter(!is.na(word_noword_levenshtein)) %>%
  count(participant_id, word_noword_levenshtein,condition, block, name = "n") %>%
  group_by(participant_id, condition, block) %>%
  mutate(proportion = n / sum(n)) %>%
  ungroup() %>%
  complete(
    participant_id,
    condition,
    block,
    word_noword_levenshtein = response_levels,
    fill = list(n = 0, proportion = 0)
  ) %>%
  mutate(word_noword_levenshtein = factor(word_noword_levenshtein, levels = response_levels))

prop_jw_cond <- dat %>%
  filter(!is.na(word_noword_jw)) %>%
  count(participant_id, word_noword_jw, condition,block, name = "n") %>%
  group_by(participant_id, condition, block) %>%
  mutate(proportion = n / sum(n)) %>%
  ungroup() %>%
  complete(
    participant_id,
    condition,
    block,
    word_noword_jw = response_levels,
    fill = list(n = 0, proportion = 0)
  ) %>%
  mutate(word_noword_jw = factor(word_noword_jw, levels = response_levels))


prop_jw_cond_midi <- subset(dat, block %in% c("mini", "midi")) %>%
  filter(!is.na(word_noword_jw)) %>%
  count(participant_id, word_noword_jw, condition, name = "n") %>%
  group_by(participant_id, condition) %>%
  mutate(proportion = n / sum(n)) %>%
  ungroup() %>%
  complete(
    participant_id,
    condition,
    word_noword_jw = response_levels,
    fill = list(n = 0, proportion = 0)
  ) %>%
  mutate(word_noword_jw = factor(word_noword_jw, levels = response_levels))

prop_jw_cond_maxi <- subset(dat, block %in% c("mini", "midi","maxi")) %>%
  filter(!is.na(word_noword_jw)) %>%
  count(participant_id, word_noword_jw, condition, name = "n") %>%
  group_by(participant_id, condition) %>%
  mutate(proportion = n / sum(n)) %>%
  ungroup() %>%
  complete(
    participant_id,
    condition,
    word_noword_jw = response_levels,
    fill = list(n = 0, proportion = 0)
  ) %>%
  mutate(word_noword_jw = factor(word_noword_jw, levels = response_levels))
```

## Plot: word_noword_levenshtein_cond

```{r}
ggplot(prop_levenshtein_cond, aes(x = word_noword_levenshtein, y = proportion, fill= condition)) +
  geom_boxplot(outlier.shape = NA, width = 0.55) +
#  geom_point(aes(group = participant_id), position = position_jitter(width = 0.12, height = 0), alpha = 0.65, size = 2) +
  scale_y_continuous(limits = c(0, 1), labels = scales::percent_format(accuracy = 1)) +
  facet_wrap(~ block,nrow=3)+
  labs(
    x = NULL,
    y = "Proportion of responses Levenshtein"
  ) +
  theme_classic(base_size = 14)
```

## Plot: word_noword_jw_cond

```{r}
ggplot(prop_jw_cond, aes(x = word_noword_jw, y = proportion, fill= condition)) +
  geom_boxplot(outlier.shape = NA, width = 0.55) +
#  geom_point(aes(group = participant_id), position = position_jitter(width = 0.12, height = 0), alpha = 0.65, size = 2) +
  scale_y_continuous(limits = c(0, 1), labels = scales::percent_format(accuracy = 1)) +
  facet_wrap(~ block, nrow=3)+
  labs(
    x = NULL,
    y = "Proportion of responses JW"
  ) +
  theme_classic(base_size = 14)
```

## Plot for correct mini, midi maxi (60, 120, 200 trials)
Calculate response proportions for the correct cumulative trial blocks (mini = 60 trials, mini + midi = 120 trials, mini + midi + maxi = 200 trials).

```{r}

a<-ggplot(subset(prop_jw_cond, block %in% c("mini")), aes(x = word_noword_jw, y = proportion, fill = condition)) +
  geom_boxplot(outlier.shape = NA, width = 0.55) +
  scale_y_continuous(
    limits = c(0, 1),
    labels = scales::percent_format(accuracy = 1)
  ) +
  #facet_wrap(~ block_plot, nrow = 3) +
  labs(
    x = NULL,
    y = "Mini: %responses"
  ) +
  theme_classic(base_size = 14)

b<-ggplot(prop_jw_cond_midi, aes(x = word_noword_jw, y = proportion, fill = condition)) +
  geom_boxplot(outlier.shape = NA, width = 0.55) +
  scale_y_continuous(
    limits = c(0, 1),
    labels = scales::percent_format(accuracy = 1)
  ) +
  #facet_wrap(~ block_plot, nrow = 3) +
  labs(
    x = NULL,
    y = "Midi: %responses"
  ) +
  theme_classic(base_size = 14)

c<-ggplot(prop_jw_cond_maxi, aes(x = word_noword_jw, y = proportion, fill = condition)) +
  geom_boxplot(outlier.shape = NA, width = 0.55) +
  scale_y_continuous(
    limits = c(0, 1),
    labels = scales::percent_format(accuracy = 1)
  ) +
  #facet_wrap(~ block_plot, nrow = 3) +
  labs(
    x = NULL,
    y = "Maxi: %responses"
  ) +
  theme_classic(base_size = 14)

grid.arrange(a,b,c,nrow=3)
```

## Calculate subject-level clarity/confidence proportions
Calculate subject-level mean clarity and confidence ratings for each condition, channel number, and block. Proportions are calculated for the correct cumulative trial blocks (mini = 60 trials, mini + midi = 120 trials, mini + midi + maxi = 200 trials). 

```{r calculate-proportions-clarity-confidence}
# clarity
prop_clarity <- dat %>%
  group_by(participant_id, condition, channels, block) %>%
  dplyr::summarize(mean_clarity = mean(clarity_rating))

prop_clarity_midi <- subset(dat, block %in% c("mini", "midi")) %>%
  group_by(participant_id, condition, channels) %>%
  dplyr::summarize(mean_clarity = mean(clarity_rating))
  
prop_clarity_maxi <- subset(dat, block %in% c("mini", "midi", "maxi")) %>%
  group_by(participant_id, condition, channels) %>%
  dplyr::summarize(mean_clarity = mean(clarity_rating))

# confidence
prop_confidence <- dat %>% 
  group_by(participant_id, condition, channels, block) %>%
  dplyr::summarize(mean_confidence = mean(confidence_rating))

prop_confidence_midi <- subset(dat, block %in% c("mini", "midi")) %>%
  group_by(participant_id, condition, channels) %>%
  dplyr::summarize(mean_confidence = mean(confidence_rating))

prop_confidence_maxi <- subset(dat, block %in% c("mini", "midi", "maxi")) %>%
  group_by(participant_id, condition, channels) %>%
  dplyr::summarize(mean_confidence = mean(confidence_rating))
```

## Plot: mean clarity
Note: Proportions are calculated separately for each block (i.e., mini = 60 trials, midi = 60 trials, maxi = 80 trials).

```{r}
ggplot(prop_clarity, aes(x = condition, y = mean_clarity, fill= channels)) +
  geom_boxplot(outlier.shape = NA, width = 0.55) +
#  geom_point(aes(group = participant_id), position = position_jitter(width = 0.12, height = 0), alpha = 0.65, size = 2) +
  scale_y_continuous(limits = c(0, 1), labels = scales::percent_format(accuracy = 1)) +
  facet_wrap(~ block, nrow=3)+
  labs(
    x = NULL,
    y = "Mean Clarity",
  ) +
  theme_classic(base_size = 14)
```

## Plot for correct clarity: mini, midi maxi (60, 120, 200 trials)
Calculate response proportions for the correct cumulative trial blocks (mini = 60 trials, mini + midi = 120 trials, mini + midi + maxi = 200 trials).

```{r}
a<-ggplot(subset(prop_clarity, block %in% c("mini")), aes(x = condition, y = mean_clarity, fill= channels)) +
  geom_boxplot(outlier.shape = NA, width = 0.55) +
  scale_y_continuous(limits = c(0, 1), labels = scales::percent_format(accuracy = 1)) +
  labs(
    x = NULL,
    y = "Mini: Mean Clarity",
  ) +
  theme_classic(base_size = 14)

b<-ggplot(prop_clarity_midi,  aes(x = condition, y = mean_clarity, fill= channels)) +
  geom_boxplot(outlier.shape = NA, width = 0.55) +
  scale_y_continuous(limits = c(0, 1), labels = scales::percent_format(accuracy = 1)) +
  labs(
    x = NULL,
    y = "Midi: Mean Clarity",
  ) +
  theme_classic(base_size = 14)

c<-ggplot(prop_clarity_maxi,  aes(x = condition, y = mean_clarity, fill= channels)) +
  geom_boxplot(outlier.shape = NA, width = 0.55) +
  scale_y_continuous(limits = c(0, 1), labels = scales::percent_format(accuracy = 1)) +
  labs(
    x = NULL,
    y = "Maxi: Mean Clarity",
  ) +
  theme_classic(base_size = 14)

grid.arrange(a,b,c,nrow=3)
```

## Plot: mean confidence
Note: Proportions are calculated separately for each block (i.e., mini = 60 trials, midi = 60 trials, maxi = 80 trials).

```{r}
ggplot(prop_confidence, aes(x = condition, y = mean_confidence, fill= channels)) +
  geom_boxplot(outlier.shape = NA, width = 0.55) +
#  geom_point(aes(group = participant_id), position = position_jitter(width = 0.12, height = 0), alpha = 0.65, size = 2) +
  scale_y_continuous(limits = c(0, 1), labels = scales::percent_format(accuracy = 1)) +
  facet_wrap(~ block, nrow=3)+
  labs(
    x = NULL,
    y = "Mean Confidence",
  ) +
  theme_classic(base_size = 14)
```

## Plot for correct confidence: mini, midi maxi (60, 120, 200 trials)
Calculate response proportions for the correct cumulative trial blocks (mini = 60 trials, mini + midi = 120 trials, mini + midi + maxi = 200 trials).

```{r}
a<-ggplot(subset(prop_confidence, block %in% c("mini")), aes(x = condition, y = mean_confidence, fill= channels)) +
  geom_boxplot(outlier.shape = NA, width = 0.55) +
  scale_y_continuous(
    limits = c(0, 1),
    labels = scales::percent_format(accuracy = 1)
  ) +
  labs(
    x = NULL,
    y = "Mini: Mean confidence"
  ) +
  theme_classic(base_size = 14)

b<-ggplot(prop_confidence_midi, aes(x = condition, y = mean_confidence, fill= channels)) +
  geom_boxplot(outlier.shape = NA, width = 0.55) +
  scale_y_continuous(
    limits = c(0, 1),
    labels = scales::percent_format(accuracy = 1)
  ) +
  labs(
    x = NULL,
    y = "Midi: Mean confidence"
  ) +
  theme_classic(base_size = 14)

c<-ggplot(prop_confidence_maxi, aes(x = condition, y = mean_confidence, fill= channels)) +
  geom_boxplot(outlier.shape = NA, width = 0.55) +
  scale_y_continuous(
    limits = c(0, 1),
    labels = scales::percent_format(accuracy = 1)
  ) +
  labs(
    x = NULL,
    y = "Maxi: Mean confidence"
  ) +
  theme_classic(base_size = 14)

grid.arrange(a,b,c,nrow=3)
```

# Create typed_word column
Correct responses ('correctword') are replaced with the according target word and are thus standardized for the cloze probability calculation.
Note: 'typed_word' is a legacy name from the (original) German version of the experiment where responses were typed. In the current version, responses are spoken and transcribed, but the column name was kept for consistency with the downstream code.  

```{r}
dat
dat$typed_word <- ifelse(
  dat$word_noword_jw == "correctword",
  dat$Answer,
  ifelse(
    dat$word_noword_jw %in% c("word", "noword"),
    dat$mic_input_text_clean,
    NA
  )
)  
```

# Create cloze_probability_typed_word
Compute cloze probability
```{r}
# create "cloze_probability_typed_word"
dat$cloze_probability_typed_word <- dat$typed_word

# comment the following line in if no word response should *not* have a cloze probability value 
#dat$cloze_probability_typed_word <- ifelse(dat$cloze_probability_typed_word == "nichts", NA, dat$cloze_probability_typed_word)

dat <- dat %>%
  group_by(Sentence, condition) %>%
  mutate(cloze_probability_typed_word_unique_predictions = n_distinct(cloze_probability_typed_word, na.rm=TRUE))

dat<- dat %>%
  group_by(Sentence, condition) %>%
  mutate(cloze_probability_typed_word_all_predictions = sum(!is.na(cloze_probability_typed_word)))

dat <- dat %>%
  group_by(Sentence, cloze_probability_typed_word) %>%
  mutate(cloze_probability_typed_word_specific = sum(!is.na(cloze_probability_typed_word)))
dat$cloze_probability_typed_word_specific[dat$cloze_probability_typed_word_specific == 0] <- NA

dat$cloze_probability_typed_word <- (dat$cloze_probability_typed_word_specific/dat$cloze_probability_typed_word_all_predictions)

#*** sanity check: check if "nichts" responses show NA for cloze_probability_typed_word ***
#*** sanity check: check if cloze_probability_typed_word sums up to 1

dat_cloze_probability_typed_word <- dat %>%
  distinct(Sentence, typed_word, cloze_probability_typed_word) %>%
  group_by(Sentence) %>%
  dplyr::summarize(cloze_prob_sum = sum(cloze_probability_typed_word, na.rm = TRUE))
#dat_cloze_probability_typed_word$cloze_prob_sum
```

# Cleanup and export
```{r}
dat <- dat[, !(names(dat) %in% c("Comments","Besonderheit","prior_expectation","expectation_confidence", "prior_input_file", "trial_index_first", "trial_index_last", "soundex", "mic_input_stt_error", "prior_input_text", "prior_input_stt_backend", "prior_input_stt_confidence", "prior_input_stt_error","prior_input_stt_language"))]
names(dat)[names(dat) == "sentence_index"] <- "trial_number"
names(dat)[names(dat) == "participant_id"] <- "participant_private_id"
names(dat)[names(dat) == "cloze"] <- "cloze_probability"
names(dat)[names(dat) == "Answer"] <- "target_word"
names(dat)[names(dat) == "clarity_rating"] <- "clarity"
names(dat)[names(dat) == "confidence_rating"] <- "confidence"


write.csv(dat, ".../preprocessing/PlangEnglish_data.csv") #your path to the preprocessing folder
```

