---
title: "Word Prediction in STPD - Experiment and Data"
author: "Chris Mathys"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
library(tidyverse)
library(corrr)
library(bayesplot)
library(here)
```

## Experiment

The setup is as follows:

1.  prior+degraded target word
2.  clarity rating (1-100, (RT possible))
3.  Type the word you heard + confidence rating for answer giving,

please see the PPT attached.

The channel numbers are 1,3,6,12. (50 sentences per channel level) -\> 200 trials in total.

The prior is continuous in entropy from very low (completely predictable) to very high (completely unpredictable)

The target words' cloze probability (expectancy) ranges from 0 (not expected) to 1 (very expected).

## Data

### Read them in

```{r}
(data_maxi <- read.csv("/Users/polina/Documents/Documents\ -\ Polina’s\ MacBook\ Air/KnolleLab/Predictive_Language_Task/IDP-main/preprocessing/PlangEnglish_data.csv") %>%
  #".../preprocessing/PlangEnglish_data.csv") #your path to the preprocessing folder 
    mutate(trial_number = as_factor(trial_number)) %>%
    mutate(participant_private_id = as_factor(participant_private_id)) %>%
    mutate(condition = as_factor(condition)) %>%
    mutate(word_noword = as_factor(word_noword_jw)))

```

### Write them out

```{r}
saveRDS(data_maxi, file = "../data/data_maxi.rds")

mini<-data_maxi%>%filter(str_detect(block, "mini"))
midi<-data_maxi%>%filter(str_detect(block, "mini|midi"))

saveRDS(midi, file = "../data/data_midi.rds")
saveRDS(mini, file = "../data/data_mini.rds")
```


### Variable overview

We have three kinds of variables:

*Outcomes (dependent variables)* 

-   `typed_word`
-   `clarity`
-   `confidence`
-   `cloze_probability_typed_word`
-   `word_noword` (Jaro-Winkler classification)

*Trial-level predictors*

-   `trial_number`
-   `channels`
-   `entropy`
-   `condition`
-   `cloze_probability`
-   `target_word`

*Participant-level predictors (variables varying between predictors)*

-   `participant_private_id`
-   `spq_{total,aeb,sanh,pi,sanx,ecc,ds}`

### Remarks
-   `condition` is defined by binning `entropy`

