---
title: "Word Prediction in STPD - Preprocessing"
author: "Chris Mathys"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_document
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
library(tidyverse)
```


### Load the data for MAXI

```{r}
(data <- readRDS("../data/data_maxi.rds"))
any(is.na(data$cloze_probability_typed_word))
data <- data[complete.cases(data$cloze_probability_typed_word), ]
```

### Preprocess

A tibble for the preprocessed data.

```{r}
ppdata <- data
```


```{r}
ppdata <- ppdata %>%
  mutate(ID = dense_rank(participant_private_id))
```


Rename variables for convenience.

```{r}
ppdata <- ppdata %>%
    rename(pt = ID) %>%
    rename(u = target_word) %>%
    rename(y = typed_word) %>%
    rename(cloze_pu = cloze_probability) %>%
    rename(cloze_py = cloze_probability_typed_word)
```

Exclude trials where no answer was given.

```{r}
ppdata <- ppdata %>% filter(word_noword != "noword")
```

Create an indicator variable for correct answers.

```{r}
ppdata <- ppdata %>% mutate(correct = as.numeric(word_noword == "correctword"))
```

Log-transform, centre, and standardize `channels`.

```{r}
ppdata <- ppdata %>%
    mutate(channels = log(channels)) %>%
    mutate(channels = channels - mean(channels)) %>%
    mutate(channels = channels/sd(channels))
```

Centre and standardize `clarity`.

```{r}
ppdata <- ppdata %>%
    mutate(clarity = clarity - mean(clarity)) %>%
    mutate(clarity = clarity/sd(clarity))
```


Centre and standardize `confidence`.

```{r}
ppdata <- ppdata %>%
    mutate(confidence = confidence - mean(confidence)) %>%
    mutate(confidence = confidence/sd(confidence))
```

Increase to a minimum of 0.01, log-transform, centre, and standardize `entropy`.

```{r}
ppdata <- ppdata %>%
    mutate(entropy = ifelse(entropy < 0.01, 0.01, entropy)) %>%
    mutate(entropy = log(entropy)) %>%
    mutate(entropy = entropy - mean(entropy)) %>%
    mutate(entropy = entropy/sd(entropy))
```

Increase to a minimum of 0.01 `cloze_pu` and `cloze_py`.

```{r}
ppdata <- ppdata %>%
     mutate(cloze_pu = ifelse(cloze_pu < 0.01, 0.01, cloze_pu)) %>%
     mutate(cloze_py = ifelse(cloze_py < 0.01, 0.01, cloze_py))
```

Centre and standardize `spq_total` in a new variable `spq`.
-> change for positive traits

### only for model 3 and 5 (only patients) ! Dont run for model 2 and 4

```{r}

ppdata <- ppdata %>%
    mutate(spq = spq_total - mean(spq_total)) %>%  
    mutate(spq = spq/sd(spq))

```


### Write the preprocessed data out

```{r}
saveRDS(ppdata, file = "../data/ppdata_maxi.rds")

```

### Load the data for MIDI

```{r}
(data <- readRDS("../data/data_midi.rds"))
any(is.na(data$cloze_probability_typed_word))
data <- data[complete.cases(data$cloze_probability_typed_word), ]
```

### Preprocess

A tibble for the preprocessed data.

```{r}
ppdata <- data
```


```{r}
ppdata <- ppdata %>%
  mutate(ID = dense_rank(participant_private_id))
```


Rename variables for convenience.

```{r}
ppdata <- ppdata %>%
    rename(pt = ID) %>%
    rename(u = target_word) %>%
    rename(y = typed_word) %>%
    rename(cloze_pu = cloze_probability) %>%
    rename(cloze_py = cloze_probability_typed_word)
```

Exclude trials where no answer was given.

```{r}
ppdata <- ppdata %>% filter(word_noword != "noword")
```

Create an indicator variable for correct answers.

```{r}
ppdata <- ppdata %>% mutate(correct = as.numeric(word_noword == "correctword"))
```

Log-transform, centre, and standardize `channels`.

```{r}
ppdata <- ppdata %>%
    mutate(channels = log(channels)) %>%
    mutate(channels = channels - mean(channels)) %>%
    mutate(channels = channels/sd(channels))
```

Centre and standardize `clarity`.

```{r}
ppdata <- ppdata %>%
    mutate(clarity = clarity - mean(clarity)) %>%
    mutate(clarity = clarity/sd(clarity))
```


Centre and standardize `confidence`.

```{r}
ppdata <- ppdata %>%
    mutate(confidence = confidence - mean(confidence)) %>%
    mutate(confidence = confidence/sd(confidence))
```

Increase to a minimum of 0.01, log-transform, centre, and standardize `entropy`.

```{r}
ppdata <- ppdata %>%
    mutate(entropy = ifelse(entropy < 0.01, 0.01, entropy)) %>%
    mutate(entropy = log(entropy)) %>%
    mutate(entropy = entropy - mean(entropy)) %>%
    mutate(entropy = entropy/sd(entropy))
```

Increase to a minimum of 0.01 `cloze_pu` and `cloze_py`.

```{r}
ppdata <- ppdata %>%
     mutate(cloze_pu = ifelse(cloze_pu < 0.01, 0.01, cloze_pu)) %>%
     mutate(cloze_py = ifelse(cloze_py < 0.01, 0.01, cloze_py))
```

Centre and standardize `spq_total` in a new variable `spq`.
-> change for positive traits

### only for model 3 and 5 (only patients) ! Dont run for model 2 and 4

```{r}

ppdata <- ppdata %>%
    mutate(spq = spq_total - mean(spq_total)) %>%  
    mutate(spq = spq/sd(spq))

```


### Write the preprocessed data out

```{r}
saveRDS(ppdata, file = "../data/ppdata_midi.rds")

```


### Load the data for MINI

```{r}
(data <- readRDS("../data/data_mini.rds"))
any(is.na(data$cloze_probability_typed_word))
data <- data[complete.cases(data$cloze_probability_typed_word), ]
```

### Preprocess

A tibble for the preprocessed data.

```{r}
ppdata <- data
```


```{r}
ppdata <- ppdata %>%
  mutate(ID = dense_rank(participant_private_id))
```


Rename variables for convenience.

```{r}
ppdata <- ppdata %>%
    rename(pt = ID) %>%
    rename(u = target_word) %>%
    rename(y = typed_word) %>%
    rename(cloze_pu = cloze_probability) %>%
    rename(cloze_py = cloze_probability_typed_word)
```

Exclude trials where no answer was given.

```{r}
ppdata <- ppdata %>% filter(word_noword != "noword")
```

Create an indicator variable for correct answers.

```{r}
ppdata <- ppdata %>% mutate(correct = as.numeric(word_noword == "correctword"))
```

Log-transform, centre, and standardize `channels`.

```{r}
ppdata <- ppdata %>%
    mutate(channels = log(channels)) %>%
    mutate(channels = channels - mean(channels)) %>%
    mutate(channels = channels/sd(channels))
```

Centre and standardize `clarity`.

```{r}
ppdata <- ppdata %>%
    mutate(clarity = clarity - mean(clarity)) %>%
    mutate(clarity = clarity/sd(clarity))
```


Centre and standardize `confidence`.

```{r}
ppdata <- ppdata %>%
    mutate(confidence = confidence - mean(confidence)) %>%
    mutate(confidence = confidence/sd(confidence))
```

Increase to a minimum of 0.01, log-transform, centre, and standardize `entropy`.

```{r}
ppdata <- ppdata %>%
    mutate(entropy = ifelse(entropy < 0.01, 0.01, entropy)) %>%
    mutate(entropy = log(entropy)) %>%
    mutate(entropy = entropy - mean(entropy)) %>%
    mutate(entropy = entropy/sd(entropy))
```

Increase to a minimum of 0.01 `cloze_pu` and `cloze_py`.

```{r}
ppdata <- ppdata %>%
     mutate(cloze_pu = ifelse(cloze_pu < 0.01, 0.01, cloze_pu)) %>%
     mutate(cloze_py = ifelse(cloze_py < 0.01, 0.01, cloze_py))
```

Centre and standardize `spq_total` in a new variable `spq`.
-> change for positive traits

### only for model 3 and 5 (only patients) ! Dont run for model 2 and 4

```{r}

ppdata <- ppdata %>%
    mutate(spq = spq_total - mean(spq_total)) %>%  
    mutate(spq = spq/sd(spq))

```


### Write the preprocessed data out

```{r}
saveRDS(ppdata, file = "../data/ppdata_mini.rds")

```



















