---
title: "Word Prediction in STPD - Model 3 (direct effects)"
author: "Chris Mathys/Franziska Knolle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
remove(list = ls())
#.rs.restartR()

library(tidyverse)
library(rstan)

options(mc.cores = parallel::detectCores())
rstan_options(auto_write = TRUE)

```


### Load the preprocessed data MAXI

```{r}
ppdata <- readRDS("../data/ppdata_maxi.rds")

```

### Set up the data structure

How many trials are there?

```{r}
(I <- ppdata %>% nrow())
```

There are 4 trial-level linear predictors (`channels`, `entropy`, `cloze_pu`, and `clarity`), plus the intercept.

```{r}
K <- 5
channels <- ppdata %>% pull(channels)
entropy <- ppdata %>% pull(entropy)
cloze_pu <- ppdata %>% pull(cloze_pu)
clarity <- ppdata %>% pull(clarity)
```

There is one participant level linear predictor (`sqp`), plus the intercept.

```{r}
L <- 2
spq <- ppdata %>% pull(spq)
```

The model also needs participant indices `pt`, the cloze probabilities of answers `cloze_py`, and the correctness `correct` of responses.

```{r}
pt <- ppdata %>% pull(pt) %>% as.numeric()
cloze_py <- ppdata %>% pull(cloze_py)
correct <- ppdata %>% pull(correct)
```

Gather everything in a list and save it.

```{r}
standata <- list(I = I,
                 K = K,
                 L = L,
                 channels = channels,
                 entropy = entropy,
                 cloze_pu = cloze_pu,
                 clarity = clarity,
                 spq = spq,
                 pt = pt,
                 cloze_py = cloze_py,
                 correct = correct)

any(is.na(standata$cloze_py))
```

```{r}
saveRDS(standata, file = "../data/standata_model_3dir_maxi.rds")


```

Fit the model.

```{r}

fit <- stan(file = "../modelling//src/word_prediction_in_stpd_model_3dir.stan",
            data = standata)
```


```{r}
print(fit)
traceplot(fit)
pairs(fit)
plot(fit)
```

Save the fitted model.

```{r}
saveRDS(fit, file = "../data/stanfit_model_3dir_maxi.rds")

```

### Load the preprocessed data MIDI

```{r}
ppdata <- readRDS("../data/ppdata_midi.rds")

```

### Set up the data structure

How many trials are there?

```{r}
(I <- ppdata %>% nrow())
```

There are 4 trial-level linear predictors (`channels`, `entropy`, `cloze_pu`, and `clarity`), plus the intercept.

```{r}
K <- 5
channels <- ppdata %>% pull(channels)
entropy <- ppdata %>% pull(entropy)
cloze_pu <- ppdata %>% pull(cloze_pu)
clarity <- ppdata %>% pull(clarity)
```

There is one participant level linear predictor (`sqp`), plus the intercept.

```{r}
L <- 2
spq <- ppdata %>% pull(spq)
```

The model also needs participant indices `pt`, the cloze probabilities of answers `cloze_py`, and the correctness `correct` of responses.

```{r}
pt <- ppdata %>% pull(pt) %>% as.numeric()
cloze_py <- ppdata %>% pull(cloze_py)
correct <- ppdata %>% pull(correct)
```

Gather everything in a list and save it.

```{r}
standata <- list(I = I,
                 K = K,
                 L = L,
                 channels = channels,
                 entropy = entropy,
                 cloze_pu = cloze_pu,
                 clarity = clarity,
                 spq = spq,
                 pt = pt,
                 cloze_py = cloze_py,
                 correct = correct)

any(is.na(standata$cloze_py))
```

```{r}
saveRDS(standata, file = "../data/standata_model_3dir_midi.rds")


```

Fit the model.

```{r}

fit <- stan(file = "../modelling/src/word_prediction_in_stpd_model_3dir.stan",
            data = standata)
```


```{r}
print(fit)
traceplot(fit)
pairs(fit)
plot(fit)
```

Save the fitted model.

```{r}
saveRDS(fit, file = "../data/stanfit_model_3dir_midi.rds")

```



### Load the preprocessed data MINI

```{r}
ppdata <- readRDS("../data/ppdata_mini.rds")

```

### Set up the data structure

How many trials are there?

```{r}
(I <- ppdata %>% nrow())
```

There are 4 trial-level linear predictors (`channels`, `entropy`, `cloze_pu`, and `clarity`), plus the intercept.

```{r}
K <- 5
channels <- ppdata %>% pull(channels)
entropy <- ppdata %>% pull(entropy)
cloze_pu <- ppdata %>% pull(cloze_pu)
clarity <- ppdata %>% pull(clarity)
```

There is one participant level linear predictor (`sqp`), plus the intercept.

```{r}
L <- 2
spq <- ppdata %>% pull(spq)
```

The model also needs participant indices `pt`, the cloze probabilities of answers `cloze_py`, and the correctness `correct` of responses.

```{r}
pt <- ppdata %>% pull(pt) %>% as.numeric()
cloze_py <- ppdata %>% pull(cloze_py)
correct <- ppdata %>% pull(correct)
```

Gather everything in a list and save it.

```{r}
standata <- list(I = I,
                 K = K,
                 L = L,
                 channels = channels,
                 entropy = entropy,
                 cloze_pu = cloze_pu,
                 clarity = clarity,
                 spq = spq,
                 pt = pt,
                 cloze_py = cloze_py,
                 correct = correct)

any(is.na(standata$cloze_py))
```

```{r}
saveRDS(standata, file = "../data/standata_model_3dir_mini.rds")


```

Fit the model.

```{r}

fit <- stan(file = "../modelling/src/word_prediction_in_stpd_model_3dir.stan",
            data = standata)
```


```{r}
print(fit)
traceplot(fit)
pairs(fit)
plot(fit)
```

Save the fitted model.

```{r}
saveRDS(fit, file = "../data/stanfit_model_3dir_mini.rds")

```

